---
title: "Bob Ross's Joy of Painting Analysis"
description: |
  This analysis examines color usage in Bob Ross's The Joy of Painting, highlighting his aesthetic choices and frequent use of a limited, serene color palette.
author: "Aditya Rao"
date: "sys.date"
categories: [blog]
format:
  html:
    toc: true
    toc_float: true
    fig-align: center
    number-sections: false
    fig-cap-location: top
    theme: solar
---
```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE)
```

```{r loading-libraries}
library(tidyverse)
library(tidytuesdayR)
library(dplyr)
library(readr)
library(ggdark)
library(tidyr)
library(ggplot2)
library(reshape2)
library(here)
library(kableExtra)
```

```{r data}
# Read the CSV file
bob_ross <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-21/bob_ross.csv')
```
```{r}
#| echo: false
# Remove the first column
bob_ross <- select(bob_ross, -1)

# Convert relevant columns to logical
bob_ross <- bob_ross |> 
  mutate(
    across(Black_Gesso:Alizarin_Crimson, as.logical)
  )
dir.create(here("data", "2023", "2023-02-21"), recursive = TRUE, showWarnings = FALSE)

# Save the data
write_csv(
  bob_ross,
  here("data", "2023", "2023-02-21", "bob_ross.csv")
)
```

```{r data cleaning}
#| echo: false
# Summarize the color usage data
color_usage <- bob_ross %>%
  select(Black_Gesso:Alizarin_Crimson) %>%
  summarise(across(everything(), ~ sum(. == TRUE, na.rm = TRUE)))

# Convert to long format
color_usage_long <- tidyr::pivot_longer(color_usage, cols = everything(), names_to = "Color", values_to = "Count")
```

## Problem Description

Bob Ross, the iconic host of The Joy of Painting, is renowned for his ability to paint tranquil and attractive landscapes in under 30 minutes. This paper examines the colors used in his paintings to determine color usage trends, with a special emphasis on the most and least often used colors. By evaluating these elements, we obtain insight into Ross' artistic tastes and the color palettes that constitute his distinct style.

The goal of this analysis is to investigate the color palette used by Bob Ross in his television series *The Joy of Painting*. By examining the frequency and combination of colors used, we aim to uncover insights into his artistic style, which helped define the serene and picturesque landscapes he painted.

## Dataset Overview

```{r}
data_description <- data.frame(
  Feature = c("Title", 
              "Season and Episode", 
              "Image Source", 
              "Color Information", 
              "Usage Frequency"),
  Description = c("The name of each painting.",
                  "The season and episode number in which the painting was featured.",
                  "Links to the images of the paintings.",
                  "Binary indicators for the presence (True/False) of 22 specific colors across the paintings, including Titanium White, Alizarin Crimson, Van Dyke Brown, etc.",
                  "Reflects the color palettes Bob Ross employed, offering insights into his artistic choices and recurring themes.")
)

# Generate the customized table for a dark theme
kable(data_description, caption = "Key Features of the Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                position = "center") %>%
  row_spec(0, background = "#333333", color = "#FFFFFF", bold = TRUE) %>%  # Header row dark background
  row_spec(1:nrow(data_description), background = "#444444", color = "#FFFFFF") %>%  # Data rows
  column_spec(1, bold = TRUE)  # Bold the first column
```


```{r plot 1}
#| echo: false
# Separate the most and least used colors
most_used_colors <- color_usage_long %>% arrange(desc(Count)) %>% head(5)
least_used_colors <- color_usage_long %>% arrange(Count) %>% head(5)

# Plot the most used colors
plot1 <- ggplot(most_used_colors, aes(x = reorder(Color, Count), y = Count)) +
  geom_col(fill = "cyan4") +
  coord_flip() +
  labs(title = "Most Used Colors in Bob Ross Paintings",
       x = "Color",
       y = "Usage Count") 
plot1 + dark_theme_bw()
```








